---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
# dynbenchmark::reformat_tribbles("README.Rmd")
```

# funkyheatmap: Generating Funky Heatmaps for Data Frames

<!-- badges: start -->
[![R-CMD-check](https://github.com/dynverse/funkyheatmap/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dynverse/funkyheatmap/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`funkyheatmap` allows generating heatmap-like visualisations for benchmark data 
frames. Funky heatmaps can be fine-tuned by providing annotations of the 
columns and rows, which allows assigning multiple palettes or geometries
or grouping rows and columns together in categories.

## Installation

You can install funkyheatmap like so:

``` r
install.packages("funkyheatmap")
```

You can also download `funkyheatmap` as a [standalone executable](http://funkyheatmap.dynverse.org/articles/executable.html) 
or a [Nextflow pipeline](http://funkyheatmap.dynverse.org/articles/nextflow.html).

## Example

Let's use the `mtcars` dataset as an example of what a funky heatmap looks like.

```{r example}
library(funkyheatmap)
library(dplyr, warn.conflicts = FALSE)
library(tibble, warn.conflicts = FALSE)

data("mtcars")
```

You can visualise the dataset as follows.

```{r include=FALSE}
g <- funky_heatmap(mtcars)
# set width and height for upcoming plot
knitr::opts_chunk$set(fig.width = g$width, fig.height = g$height)
```

```{r eval=FALSE}
funky_heatmap(mtcars)
```

```{r heatmap1, echo=FALSE}
g
```

## Customising the plot
However, this plot can look so much better if you provide additional metadata
for the rows and columns. See more information on how you can customise.

```{r example-with-colinfo}
data <- mtcars %>%
  rownames_to_column("id") %>%
  arrange(desc(mpg))

column_info <- tribble(
  ~id,     ~group,         ~name,                      ~geom,        ~palette,    ~options,
  "id",    "",             "",                         "text",       NA,          list(hjust = 0, width = 6),
  "mpg",   "overall",      "Miles / gallon",           "bar",        "palette1",  list(width = 4, legend = FALSE),
  "cyl",   "overall",      "Number of cylinders",      "bar",        "palette2",  list(width = 4, legend = FALSE),
  "disp",  "group1",       "Displacement (cu.in.)",    "funkyrect",  "palette1",  lst(),
  "hp",    "group1",       "Gross horsepower",         "funkyrect",  "palette1",  lst(),
  "drat",  "group1",       "Rear axle ratio",          "funkyrect",  "palette1",  lst(),
  "wt",    "group1",       "Weight (1000 lbs)",        "funkyrect",  "palette1",  lst(),
  "qsec",  "group2",       "1/4 mile time",            "circle",     "palette2",  lst(),
  "vs",    "group2",       "Engine",                   "circle",     "palette2",  lst(),
  "am",    "group2",       "Transmission",             "circle",     "palette2",  lst(),
  "gear",  "group2",       "# Forward gears",          "circle",     "palette2",  lst(),
  "carb",  "group2",       "# Carburetors",            "circle",     "palette2",  lst()
)
```

```{r include=FALSE}
g <- funky_heatmap(data, column_info = column_info, expand = list(xmax = 4))
knitr::opts_chunk$set(fig.width = g$width, fig.height = g$height)
```

```{r eval=FALSE}
funky_heatmap(data, column_info = column_info, expand = list(xmax = 4))
```

```{r heatmap2, echo=FALSE}
g
```

## More information 
  
* The [reference documentation](http://funkyheatmap.dynverse.org/reference/index.html)
  on `funky_heatmap()` details the exact formats of each annotation object that you can pass to it.

* Check out the vignette [`vignette("mtcars", "funkyheatmap")`](http://funkyheatmap.dynverse.org/articles/mtcars.html)
  for more information on how to customize this visualisation.

* In [`vignette("dynbenchmark", "funkyheatmap")`](http://funkyheatmap.dynverse.org/articles/dynbenchmark.html) we 
  use funkyheatmap to regenerate the figures from Saelens et al. (2019)
  [doi:10.1038/s41587-019-0071-9](https://doi.org/10.1038/s41587-019-0071-9).

* We used [Viash](https://viash.io) to wrap the `funkyheatmap::funky_heatmap()` function
  as a [standalone executable](http://funkyheatmap.dynverse.org/articles/executable.html) and
  [Nextflow module](http://funkyheatmap.dynverse.org/articles/nextflow.html).
