---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# funkyheatmap: Generating Funky Heatmaps for Data Frames

<!-- badges: start -->
[![R-CMD-check](https://github.com/rcannood/funkyheatmap/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rcannood/funkyheatmap/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Allows generating beautiful visualisations for benchmark data 
frames. Funky heatmaps can be fine-tuned by providing annotations of the 
columns and rows, which allows assigning multiple palettes or geometries
or grouping rows and columns together in categories.

## Installation

You can install the development version of funkyheatmap like so:

``` r
devtools::install_github("rcannood/funkyheatmap")
```

## Example

You can visualise the `mtcars` dataset as follows.

```{r example}
library(funkyheatmap)
library(tibble)
library(purrr)

data("mtcars")

data <- mtcars %>% rownames_to_column("id")

g <- funky_heatmap(data)
```

```{r include=FALSE}
# set width and height for upcoming plot
knitr::opts_chunk$set(fig.width = g$width, fig.height = g$height)
```

```{r example-heatmap}
g
```

However, this plot can look so much better if you provide additional metadata
for the rows and columns.


## More information 

* Check out the vignette `vignette("mtcars", "funkyheatmap")` for more 
  information on how to customize this visualisation.

* The vignette `vignette("dynbenchmark", "funkyheatmap")`  displays the code 
  that was used in Saelens et al. (2019) [doi:10.1038/s41587-019-0071-9](https://doi.org/10.1038/s41587-019-0071-9).
  
* The reference documentation on `funky_heatmap()` details the exact formats of 
  each annotation object that you can pass to it.
