---
title: "Getting started with funkyheatmap"
vignette: >
  %\VignetteIndexEntry{Getting started with funkyheatmap"}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  fig.width = 8,
  fig.height = 11.5
)

```

We will use the mtcars dataset to showcase how you can use **{funkyheatmap}** to visualise a dataset. We will gradually add features to 
the plot to show how you can customise it to your liking.

```{r check_on_cran, include=FALSE}
on_cran <- !identical(Sys.getenv("NOT_CRAN"), "true")

if (on_cran) {
  knitr::asis_output("Note: The outputs of this vignette are not rendered on CRAN due to package size limitations. Please check the [Getting started](https://funkyheatmap.github.io/funkyheatmap/articles/funkyheatmap.html) vignette in the package documentation.")
} else {
  knitr::opts_chunk$set(eval = FALSE)
}
```

## Loading the data

First, we load in the data. We move the rownames to a column named `id`, as that is required by **{funkyheatmap}** to display the names.
Next, we sort the dataframe on the `mpg` column in descending order, and select the top 30 rows.

```{r load-data}
library(funkyheatmap)
library(dplyr, warn.conflicts = FALSE)
library(tibble, warn.conflicts = FALSE)
library(purrr, warn.conflicts = FALSE)

data("mtcars")

data <- mtcars %>%
  rownames_to_column("id") %>%
  arrange(desc(mpg)) %>%
  head(30)
```

We can plot this data frame without any additional formatting, though it doesn't look very nice:
```{r basic-fh}
funky_heatmap(data)
```

## Adding column_info

We can add column information to the heatmap by specifying the `column_info` argument. 
This argument should be a dataframe or a tibble, with each row corresponding to a column in the heatmap.
The dataframe should have the following column:
- `id (character)`: the name of the column in the heatmap

Other columns are optional and will determine the appearance of the heatmap.
If you want to group some of the columns together, usually because they are semantically related,
or another logical grouping exists, we need to sepficy the `group` column in the `column_info` dataframe.

```{r column-group}
cinfo <- tibble(
  id = colnames(data),
  group = c(NA, "Overall", "Engine", "Engine", "Engine", "Transmission", "Overall", "Performance", "Engine", "Transmission", "Transmission", "Engine"),
  options = lapply(seq(12), function(x) lst())
)
cinfo
```

```{r column-info-fh}
funky_heatmap(data, column_info = cinfo)
```

We can see that this does not work: we need to manually rearrange the columns in the `column_info` dataframe
so that the columns in the same column groups are adjacent to each other.

```{r column-group-sort}
data <- data[, c("id", "qsec", "mpg", "wt", "cyl", "carb", "disp", "hp", "vs", "drat", "am", "gear")]

cinfo <- tibble(
  id = colnames(data),
  group = c(NA, "Performance", rep("Overall", 2), rep("Engine", 5), rep("Transmission", 3)),
  options = lapply(seq(12), function(x) lst())
)
cinfo

funky_heatmap(data, column_info = cinfo)
```

We should probably also add a `name` column, to make the column labels more informative.

```{r column-info-name}
cinfo$name <- c("Model", "1/4 mile time", "Miles per gallon", "Weight", "Number of cylinders", "Carburetors", "Displacement", "Horsepower", "Engine type", "Rear axle ratio", "Transmission", "Forward gears")
funky_heatmap(data, column_info = cinfo)
```

This is better, we can at least read the column groups now. However, the column groups are not very clear: 
we can add a `palette` column to the `column_info` dataframe to make the column groups more visually distinct.

## Adding palettes

Just adding a `palette` column is not good enough: we need to specify a separate palette list as well.
The easiest way to do this is by creating a named list of palettes, where the names correspond to the values 
in the `palette` column of the `column_info` dataframe and use one of the predefined palettes. These are:

for numerical data:
- `Blues`
- `Reds`
- `Greens`
- `YlOrBr`
- `Greys`

and for categorical data:
- `Set1`
- `Set2`
- `Set3`
- `Dark2`

```{r column-info-palette}
cinfo$palette <- c(NA, "perf_palette", rep("overall_palette", 2), rep("engine_palette", 5), rep("transmission_palette", 3))

palettes <- list(perf_palette = "Blues", overall_palette = "Greens", engine_palette = "YlOrBr", transmission_palette = "Reds")

funky_heatmap(data, column_info = cinfo, palettes = palettes)
```

## Adding column_groups

That's already a lot more visually distinct! It would be nice if we could color the names of the column groups as well.
We can do that by adding a `column_group` tibble or dataframe.
It must have the following columns:
- `Category (character)`: the name of the column group that will be displayd
- `group (character)`: the group that the column group belongs to. This should correspond to the group label in the `column_info` dataframe
- `palette (character)`: the palette to use for the column group. 

```{r column-group-df}
column_groups <- tibble(
  Category = c("Performance", "Overall", "Engine", "Transmission"),
  group = c("Performance", "Overall", "Engine", "Transmission"),
  palette = c("perf_palette", "overall_palette", "engine_palette", "transmission_palette")
)

funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes)
```

## Specifying geoms
It makes sense for some of the information to be presented in a different way.
For instance, the number of cylinders and the number of carburetors are discrete values, so it would make sense to display them as 
rectangles with a text overlay.

```{r column-info-geom}
cinfo$geom <- c("text", "bar", "bar", "bar", "rect", "rect", "funkyrect", "funkyrect", "circle", "funkyrect", "rect", "rect")
funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes)
```

This looks nice, but we don't have the text overlay yet.
```{r column-info-text}
# column_info$options <- lapply(seq(12), function(x) lst())
cinfo <- cinfo %>% 
  add_row(id = "cyl", group = "Engine", name = "", geom = "text", options = lst(lst(overlay = TRUE)), palette = "black", .before = 6) %>%
  add_row(id = "carb", group = "Engine", name = "", geom = "text", options = lst(lst(overlay = TRUE)), palette = "black", .before = 8) %>%
  add_row(id = "am", group = "Transmission", name = "", geom = "text", options = lst(lst(overlay = TRUE)), palette = "black", .before = 14) %>%
  add_row(id = "gear", group = "Transmission", name = "", geom = "text", options = lst(lst(overlay = TRUE)), palette = "black", .before = 17)

cinfo

palettes$black <- c(rep("black", 2))
funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes)
```

This is starting to look like a nice visualisation!

## Customising legends

We can also customise the legends for the different columns.
If we want multiple legends for a single palette, we just need to specify multiple entries in the `legends` list
for that palette.

```{r legends}
palettes$funky_palette_grey <- RColorBrewer::brewer.pal(9, "Greys")[-1] %>% rev()

legends <- list(
    list(
        palette = "perf_palette",
        geom = "bar",
        title = "1/4 mile time",
        labels = c(paste0(min(data$qsec), "s"), rep("", 8), paste0(max(data$qsec), "s"))
    ),
    list(
        palette = "overall_palette",
        geom = "bar",
        title = "Miles per gallon",
        labels = c(paste0(min(data$mpg), "mpg"), rep("", 8), paste0(max(data$mpg), "mpg"))
    ),
    list(
        palette = "overall_palette",
        geom = "bar",
        title = "Weight",
        labels = c(paste0(min(data$wt), "lbs"), rep("", 8), paste0(max(data$wt), "lbs"))
    ),
    list(
        palette = "funky_palette_grey",
        geom = "funkyrect",
        title = "Overall",
        enabled = TRUE,
        labels = c("0", "", "0.2", "", "0.4", "", "0.6", "", "0.8", "", "1")
    )
)
funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes, legends = legends)
```

Now, we would like to disable the `engine_palette` and the `transmission_palette` legends, as they don't really add extra information anymore.
We can do this with the `enabled` argument in `legends` list.

```{r legends-disable}
disabled_legends = list(
    list(
        palette = "engine_palette",
        enabled = FALSE
    ),
    list(
        palette = "transmission_palette",
        enabled = FALSE
    )
)

# append disabled_legends to legends
legends <- c(legends, disabled_legends)

funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes, legends = legends)
```


## Adding images

The `transmission` and `engine type` columns are still not very informative: they basically just contain 0 or 1, with no information what this means.
We can replace these values with images to make it more clear.
If we want to replace their values with images, we need to specify the image locations.

First we replace their values by the image names in the actual data.
We will specify the directory and extension in the `column_info` dataframe.

We also need to change the `geom` in `column_info` to `image` for the columns that should display images.

```{r images}
# change the am: if 0 go to "automatic", if 1 go to "manual"
data[data$am == 0, "am"] <- "automatic"
data[data$am == 1, "am"] <- "manual"

# change the vs: if 0 go to "vengine", if 1 go to "straight"
data[data$vs == 0, "vs"] <- "vengine"
data[data$vs == 1, "vs"] <- "straight"

cinfo$directory <- NA
cinfo$extension <- NA

# remove row 14
cinfo <- cinfo[-14, ]

cinfo[cinfo$id %in% c("vs", "am"), "directory"] <- "images"
cinfo[cinfo$id %in% c("vs", "am"), "extension"] <- "png"
cinfo[c(11, 13), "geom"] <- "image"

funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes, legends = legends)
```


## Row grouping

We can also group rows together. Let's say that in this case we especially want to highlight the Mercedes cars.
For this, we need to rearrange the data so that the Mercedes cars are at the top of the dataframe.
We can then add a `row_info` dataframe that specifies the grouping of the rows.

```{r row-grouping}
row_info <- data %>% transmute(id, group = ifelse(grepl("Merc", id), "Mercedes", "Other"))
# sort Mercedes cars to the top of the data and the row_info dataframe
data <- data[order(row_info$group), ]
row_info <- row_info[order(row_info$group), ]

row_groups <- tibble(level1 = c("Mercedes", "Other cars"), group = c("Mercedes", "Other"))

funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes, legends = legends, row_info = row_info, row_groups = row_groups)
```

Now, we can perform some small tweaks with regard to spacing: e.g. the `Transmission` column group needs slightly more space.
We can do this by specifying the `width` argument in the `options` list in the `column_info` dataframe.

If there are still some small things you want to tweak, it is recommended to save this plot as an `.svg` file 
and use a vector based graphics editor to make some final tweaks.

```{r additions}
# set options of performance column
cinfo[[1, "options"]] <- list(list(width = 6))
cinfo[[2, "options"]] <- list(list(width = 6))
cinfo[[3, "options"]] <- list(list(width = 3))
cinfo[[4, "options"]] <- list(list(width = 3))
cinfo[[12, "options"]] <- list(list(width = 1.85))
cinfo[[13, "options"]] <- list(list(width = 1.85))

funky_heatmap(data, column_info = cinfo, column_groups = column_groups, palettes = palettes, legends = legends, row_info = row_info, row_groups = row_groups)

```
