---
title: "Small funky heatmap using mtcars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Small funky heatmap using mtcars}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# rerun this to update tribble formatting:
# dynbenchmark::reformat_tribbles("vignettes/mtcars.Rmd")
```

Load data
```{r load-data}
library(funkyheatmap)
library(dplyr, warn.conflicts = FALSE)
library(tibble, warn.conflicts = FALSE)
library(purrr, warn.conflicts = FALSE)

data("mtcars")

data <- mtcars %>% 
  rownames_to_column("id") %>%
  arrange(desc(mpg)) %>%
  head(20)
```

We can plot this data frame without any additional formatting, though it doesn't look very nice:
```{r}
funky_heatmap(data)
```

By defining a few additional formatting parameters, we can get the plot to look much nicer.

## Column info 


```{r column-info}
column_info <- tribble(
  ~id,     ~group,         ~name,                      ~geom,        ~palette,    ~options,
  "id",    "",             "",                         "text",       NA,          list(hjust = 0, width = 6),
  "mpg",   "overall",      "Miles / gallon",           "bar",        "palette1",  list(width = 4, legend = FALSE),
  "cyl",   "overall",      "Number of cylinders",      "bar",        "palette2",  list(width = 4, legend = FALSE),
  "disp",  "group1",       "Displacement (cu.in.)",    "funkyrect",  "palette1",  lst(),
  "hp",    "group1",       "Gross horsepower",         "funkyrect",  "palette1",  lst(),
  "drat",  "group1",       "Rear axle ratio",          "funkyrect",  "palette1",  lst(),
  "wt",    "group1",       "Weight (1000 lbs)",        "funkyrect",  "palette1",  lst(),
  "qsec",  "group2",       "1/4 mile time",            "circle",     "palette2",  lst(),
  "vs",    "group2",       "Engine",                   "circle",     "palette2",  lst(),
  "am",    "group2",       "Transmission",             "circle",     "palette2",  lst(),
  "gear",  "group2",       "# Forward gears",          "circle",     "palette2",  lst(),
  "carb",  "group2",       "# Carburetors",            "circle",     "palette2",  lst()
)
```

Define column groups
```{r column-groups}
column_groups <- tribble( # tribble_start
  ~Category,  ~group,         ~palette,
  "\n",       "method_info",  "overall",
  "Overall",  "overall",      "overall",
  "Group 1",  "group1",       "palette1",
  "Group 2",  "group2",       "palette2"
) # tribble_end
```

## Row info


Determine method grouping
```{r method-grouping}
row_info <- data %>% transmute(id, group = "test")
row_groups <- tibble(Group = "Test", group = "test")
```


## Palettes


Determine palettes
```{r column-grouping}
palettes <- tribble(
  ~palette,             ~colours,
  "overall",            grDevices::colorRampPalette(rev(RColorBrewer::brewer.pal(9, "Greys")[-1]))(101),
  "palette1",           grDevices::colorRampPalette(rev(RColorBrewer::brewer.pal(9, "Blues") %>% c("#011636")))(101),
  "palette2",           grDevices::colorRampPalette(rev(RColorBrewer::brewer.pal(9, "Reds")[-8:-9]))(101)
)
```

## Funky heatmap

Generate funky heatmap
```{r make-plot, fig.width=4.5, fig.height=10}
g <- funky_heatmap(
  data = data,
  column_info = column_info,
  column_groups = column_groups,
  row_info = row_info,
  row_groups = row_groups,
  palettes = palettes,
  expand = list(xmax = 4)
)
```

```{r include=FALSE}
# set width and height for upcoming plot
knitr::opts_chunk$set(fig.width = g$width, fig.height = g$height)
```

```{r heatmap}
g
```

`funkyheatmap` automatically recommends a width and height for the generated plot.
To save your plot, run:

``` r
ggsave("path_to_plot.pdf", g, device = cairo_pdf, width = g$width, height = g$height)
```

